Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2023-12-30 main.tex
remove e-rewriting
Refined parent and consistency relations WIP
Pushout for EHyp with consistency and proofs
TODOs added
introduction revision

% quiver is for quiver figures
% \usepackage{quiver}
\usepackage{tikz}
\usepackage{tikz-network}
\usepackage{xspace}
\usepackage{xargs}
\usepackage[
  colorinlistoftodos,
  prependcaption,
  % disable,
]{todonotes}
\usepackage{fontenc}

@@

% category names
\newcommand{\Ecospans}{{\catname{ECsp_{D}(EHyp_{\Sigma})}}}
\newcommand{\MdaEcospans}{{\catname{MAECsp_{D}(EHyp_{\Sigma})}}}
\newcommand{\WellTypedMdaEcospans}{{\catname{\overline{MAECsp_{D}(EHyp_{\Sigma})}}}}

% There is already a remark in the preamble, but it does
% not work for some reason
\newtheorem{remark}{Remark}

@@

\title{Equivalence Hypergraphs:\\ Semantics of E-Graphs and their Generalization to 
Monoidal Theories}
\author{Chris Barrett, Dan Ghica, Aleksei Tiurin}

@@

\begin{abstract}
Abstract.
\end{abstract}

@@

\input{sections/introduction}
\input{sections/categorical_semantics}
\input{sections/combinatorial_semantics}
\input{sections/soundness_and_completeness}
\input{sections/e-graph-translation}
\input{sections/conclusion}

\subsection*{Acknowledgements}
Sadrzadeh Mehrnoosh

\clearpage
\bibliographystyle{acm}
\bibliography{bibliography}

>>>>

% quiver is for quiver figures
% including quiver redefines \Bbbk which leads to an error
% \usepackage{quiver}

\usepackage{tikz}
\usepackage{tikz-network}
\usepackage{xspace}
\usepackage{xargs}
\usepackage[
  colorinlistoftodos,
  prependcaption,
  % disable,
]{todonotes}
\usepackage{fontenc}

@@

% category names
\newcommand{\MdaCospans}{{\catname{MACsp_{D}(Hyp_{\Sigma})}}}

\newcommand{\Ecospans}{{\catname{ECsp_{D}(EHyp_{\Sigma})}}}
\newcommand{\MdaEcospans}{{\catname{MAECsp_{D}(EHyp_{\Sigma})}}}
\newcommand{\WellTypedMdaEcospans}{{\catname{\overline{MAECsp_{D}(EHyp_{\Sigma})}}}}

% conflict
\newcommand{\hashtag}{{\#}}
\newcommand{\consistency}{{\smile}}
% There is already a remark in the preamble, but it does
% not work for some reason
\newtheorem{remark}{Remark}

@@

\title{Equivalence Hypergraphs: E-Graphs for Monoidal Theories}
\author{Chris Barrett}
\author{Dan R. Ghica}
\author{Aleksei Tiurin}

@@

\begin{abstract}
The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
We give a categorical semantics for \emph{e-graphs} over algebraic theories, a generalisation of the structures which are the subject of this technique, and to their rewriting. 
The proposed framework consists of Cartesian categories enriched over a semilattice,  further generalised to symmetric monoidal categories, to account for monoidal theories.
The aim is to permit the optimisation of a larger class of programs, based on string diagrams.  
Furthermore, we present a sound and complete combinatorial representation of morphisms in such a category, based on a generalisation of hypergraphs,  which we call \emph{e-hypergraphs}, which have the usual advantage that many structural equations are absorbed into a general notion of isomorphism. 
\end{abstract}

@@

\input{sections/introduction}
\input{sections/categorical_semantics}
\input{sections/combinatorial_semantics}
\input{sections/soundness_and_completeness}
\input{sections/conclusion}

\subsection*{Acknowledgements}
We are grateful to Sadrzadeh Mehrnoosh for enlightening discussions on potential applications of e-hypergraphs to computational linguistics.

\bibliographystyle{acm}
\bibliography{bibliography}
